Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph for Community Detection and Parameter Estimation

### 2021-06-16 summary.tex
minor updates

\emph{Case \(k \neq l\)}. \(P^{(kl)}_n\) and \(A^{(kl)}_n\) represent edge
probabilities and edges between communities \(k\) and \(l\). Note that
\(P^{(kl)} = (P^{(lk)})^\top\).\\
By definition, \(P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top\). As
in the \(k = l\) case, we note that the singular value decomposition
\(P^{(kl)} = \sigma_{kl}^2 u^{(kl)} (v^{(kl)})^\top\) is one-dimensional
and \(\lambda^{(kl)} = \sigma_{kl} u^{(kl)}\). (We can also note that
the SVD of \(P^{(lk)} = \sigma_{kl}^2 v^{(kl)} (u^{(kl)})^\top\), i.e.,
\(\sigma_{kl} = \sigma_{lk}\), \(u^{(kl)} = v^{(lk)}\), and
\(v^{(kl)} = u^{(lk)}\).)\\
Now consider the Hermitian dilation

\[M^{(kl)} = 2 \begin{bmatrix} 0 & P^{(kl)} \\ P^{(lk)} & 0 \end{bmatrix}\]

which is a symmetric \(n_{kl} \times n_{kl}\) matrix. It can
be shown that the spectral decomposition of \(M^{(kl)}\) is

\[M^{(kl)} =
\begin{bmatrix} u^{(kl)} & -u^{(kl)} \\ v^{(kl)} & v^{(kl)} \end{bmatrix} \times
\begin{bmatrix} \sigma^2_{kl} & 0 \\ 0 & -\sigma^2_{kl} \end{bmatrix} \times
\begin{bmatrix} u^{(kl)} & -u^{(kl)} \\ v^{(kl)} & v^{(kl)} \end{bmatrix}^\top\]

@@

\[\begin{bmatrix}
  \sigma_{kl} u^{(kl)} & \sigma_{kl} u^{(kl)} \\
  \sigma_{kl} v^{(kl)} & -\sigma_{kl} v^{(kl)}
\end{bmatrix} =
\begin{bmatrix}
  \lambda^{(kl)} & \lambda^{(kl)} \\
  \lambda^{(lk)} & -\lambda^{(lk)}
\end{bmatrix}\]

@@

\[\hat{M}^{(kl)} = \begin{bmatrix} 0 & A^{(kl)} \\ A^{(lk)} & 0 \end{bmatrix}\]
Then \(\hat{M}^{(kl)}\) is an adjacency matrix drawn from \(M^{(kl)}\),
so its adjacency spectral embedding, given by

>>>>

\emph{Case \(k \neq l\)}. \(P^{(kl)}_n\) and \(A^{(kl)}_n\) represent edge
probabilities and edges between communities \(k\) and \(l\). Note that
\(P^{(kl)}_n = (P^{(lk)}_n)^\top\).
By definition, \(P^{(kl)}_n = \lambda^{(kl)}_n (\lambda^{(lk)}_n)^\top\).
Therefore,
\(P^{(kl)} = (\sigma^{(kl)}_n)^2 u^{(kl)}_n (v^{(kl)}_n)^\top\) is rank 2

Now consider the Hermitian dilation

\[M^{(kl)}_n = 2 \begin{bmatrix}
0 & P^{(kl)}_n \\ P^{(lk)}_n & 0 \end{bmatrix}\]

which is a symmetric \(n_{kl} \times n_{kl}\) matrix. It can
be shown that the spectral decomposition of \(M^{(kl)}_n\) is

\[M^{(kl)}_n =
\begin{bmatrix} u^{(kl)} & -u^{(kl)} \\ v^{(kl)} & v^{(kl)} \end{bmatrix} \times
\begin{bmatrix} \sigma^2_{kl} & 0 \\ 0 & -\sigma^2_{kl} \end{bmatrix} \times
\begin{bmatrix} u^{(kl)} & -u^{(kl)} \\ v^{(kl)} & v^{(kl)} \end{bmatrix}^\top\]

@@

\[\begin{bmatrix}
  \lambda^{(kl)}_n & \lambda^{(kl)}_n \\
  \lambda^{(lk)}_n & -\lambda^{(lk)}_n
\end{bmatrix}\]

@@

\[\hat{M}^{(kl)}_n = \begin{bmatrix} 0 & A^{(kl)}_n \\ A^{(lk)}_n & 0 \end{bmatrix}\]
Then \(\hat{M}^{(kl)}_n\) is an adjacency matrix drawn from \(M^{(kl)}_n\),
so its adjacency spectral embedding, given by
